---
title: "Joel's reverse engineering project notebook"
output: html_notebook
---

Loading libraries

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

Open the data

```{r}
read_csv("fatal-police-shootings-data.csv")
fatal_police_shootings <- read.csv("fatal-police-shootings-data.csv")
women_fatal_police_shootings <- fatal_police_shootings %>%  
  filter(str_detect(gender, "female")) 

count_women_police_shootings <- women_fatal_police_shootings %>% 
  summarize(count = n())

## cleaning the column names so I won't have trouble later on
clean_women_fatal_police_shootings <- women_fatal_police_shootings %>% 
  clean_names()
```

Now, **clean_women_fatal_police_shootings** represents **women_fatal_police_shootings** but cleaned to remove spaces and the like so we can actually work with column names if necessary

Next, we want to clean the "date" column to actually work like a date column with lubridate. To not mess things up with the original data, we're going to create a new variable that represents the above data but only only until 2020, when the Washington Post article was published.

```{r}
clean_2020_women_fatal_police_shootings <- clean_women_fatal_police_shootings
```

Then, we're going to actually transform the column into something useful with lubridate.

```{r}
clean_2020_women_fatal_police_shootings <- clean_2020_women_fatal_police_shootings %>% 
  mutate(date=mdy(date))
```

Now, we're finally going to actually transform it into before the publish date.

```{r}
clean_2020_women_fatal_police_shootings <- clean_2020_women_fatal_police_shootings %>% 
  filter(date<"2020-09-04") %>% 
  summarize(count = n())

clean_2020_women_fatal_police_shootings
  
```

We're getting 248 - one more than when the article was published. So we're going to try to look at the dataset as it was when the Post published its article.

```{r}
OLD_fatal_police_shootings <- read_csv("2020_sept_3_fatal-police_shootings_data.csv")
OLD_women_fatal_police_shootings <- OLD_fatal_police_shootings %>%  
  filter(str_detect(gender, "female")) 

## cleaning the column names so I won't have trouble later on
OLD_clean_count_women_fatal_police_shootings <- OLD_count_women_fatal_police_shootings %>% 
  clean_names()

class(OLD_fatal_police_shootings$date)

OLD_clean_count_women_fatal_police_shootings <- OLD_clean_count_women_fatal_police_shootings %>% 
  filter(date<"2020-09-04") %>% 
  summarize(count = n())

OLD_clean_count_women_fatal_police_shootings
```
