---
title: "Reverse Engineering Project"
author: "Taylor Joel Shaun & Menna"
date: "7/7/23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been shot by police since 2015](https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# This code sets the working directory to my local one (Taylor) - it will mess up your code, so erase this section and make sure yours is set to the correct working directory
setwd("~/data_journalism/data_journalism_fall_2023/major_assignments/reverse_engineering/Reverse-Engineering-Project/data")

# Load required data
fatal_police_shootings <- read_csv("fatal-police-shootings-data_v1.csv")
data_to_sept_2020_shootings <- read_csv("2020_sept_3_fatal-police_shootings_data.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
clean_fatal_police_shootings <- fatal_police_shootings |> 
  clean_names()

data_to_sept_2020_shootings <- data_to_sept_2020_shootings |> 
  clean_names()

women_fatal_police_shootings_sept_2020 <- data_to_sept_2020_shootings |>  
  filter(gender == "F") 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "Since The Washington Post began tracking fatal shootings by police in 2015, officers have fatally shot 247 women out of the more than 5,600 people killed overall."
* **Analysis summary**: We were able to get pretty close to the numbers quoted in the article. We came up with 248, one more than the Post included in their article. We aren't sure where the extra person is coming from, but we weren't able to locate them. The data doesn't include timing, so maybe the article came out before one from the same day got logged? That seems like a stretch though.

```{r}
# Put code to reverse engineer sentence here
women_fatal_police_shootings_sept_2020 <- women_fatal_police_shootings_sept_2020 |>
  filter(date<"2020-09-04")

women_fatal_police_shootings_sept_2020 |> 
  summarize(count = n())

# Display results of code below this codeblock

```

```{r}

fatal_police_shootings_2020 <- fatal_police_shootings_2020 |>
    filter(date<"2020-09-04")

fatal_police_shootings_2020 |> group_by(gender) |>
  summarise(count = n())

```

### Sentence 2

* **Sentence text**: Crenshaw said Black women’s deaths also may be dismissed as “collateral damage” if they are killed while police are pursuing someone else. Twenty of the 247 women were killed in that kind of situation, analysis shows. In 12 of those 20 shootings, police said the women killed were caught in crossfire or shot accidentally.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
women_fatal_police_shootings_2020 |> 
  group_by(threat_level, flee) |>
  summarise(count = n())

# using the V1 data, we aren't seeing any of the numbers we're looking for (20 and 12). The wording is kind of vague, maybe they are only talking about Black women. Let's try to filter for race.

```
```{r}
women_fatal_police_shootings_2020 |> 
  filter(race == "B")|>
   group_by(threat_level, flee) |>
  summarise(count = n())

#we still aren't seeing the same numbers. Maybe we need to look at v2 data?
```

### Sentence 3

* **Sentence text**: At least 89 of the women [fatally shot by police] were at their homes or residences where they sometimes stayed.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
fatal_police_shootings_2020 |>
  group_by(gender, was_mental_illness_related) |>
  summarise(count = n()) |>
  arrange(desc(count))

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-