---
title: "Reverse Engineering Project"
author: "Taylor Joel Shaun & Menna"
date: "7/7/23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been shot by police since 2015](https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# This code sets the working directory to my local one (Taylor) - it will mess up your code, so erase this section and make sure yours is set to the correct working directory
setwd("~/data_journalism/data_journalism_fall_2023/major_assignments/reverse_engineering/Reverse-Engineering-Project/data")

# Load required data
fatal_police_shootings <- read_csv("fatal-police-shootings-data.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
clean_fatal_police_shootings <- fatal_police_shootings |> 
  clean_names()

women_fatal_police_shootings <- clean_fatal_police_shootings |>  
  filter(gender == "female") 

clean_fatal_police_shootings <- clean_fatal_police_shootings |> mutate(date=mdy(date))
women_fatal_police_shootings <- women_fatal_police_shootings |> mutate(date=mdy(date))

clean_fatal_police_shootings_2020 <- clean_fatal_police_shootings |>  
  filter(date<"2020-09-04")
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "Since The Washington Post began tracking fatal shootings by police in 2015, officers have fatally shot 247 women out of the more than 5,600 people killed overall."
* **Analysis summary**: We were able to get pretty close to the numbers quoted in the article. We came up with 248, one more than the Post included in their article. We aren't sure where the extra person is coming from, but we weren't able to locate them. The data doesn't include timing, so maybe the article came out before one from the same day got logged? That seems like a stretch though.

```{r}
# Put code to reverse engineer sentence here
women_fatal_police_shootings_2020 <- women_fatal_police_shootings |>
  filter(date<"2020-09-04")

women_fatal_police_shootings_2020 |> 
  summarize(count = n())

# Display results of code below this codeblock

```

We're getting 1 more person than the Post did. Let's look at their original dataset to see if that makes a difference.

```{r}
setwd("~/data_journalism/data_journalism_fall_2023/major_assignments/reverse_engineering/Reverse-Engineering-Project/data")

data_to_sept_2020_shootings <- read_csv("2020_sept_3_fatal-police_shootings_data.csv")

# Clean required data and prepare for analysis if needed. 

data_to_sept_2020_shootings <- data_to_sept_2020_shootings |> 
  clean_names()

women_fatal_police_shootings_sept_2020 <- data_to_sept_2020_shootings |>  
  filter(gender == "F") 

women_fatal_police_shootings_sept_2020 <- women_fatal_police_shootings_sept_2020 |>
  filter(date<"2020-09-04")

women_fatal_police_shootings_sept_2020 |> 
  summarize(count = n())

```

It worked!

### Sentence 2

* **Sentence text**: Twenty of the 247 women were killed in that kind of situation, analysis shows. In 12 of those 20 shootings, police said the women killed were caught in crossfire or shot accidentally.
* **Analysis summary**: We weren't able to figure out where the Post got its numbers from. For one, the terms they used were quite vague (i.e. giving data about all women - 247 - while quoting Crenshaw about Black women). Not to be daunted, we tried every possibility we could think of to try to get matching numbers, including using the old dataset. Nothing brought us to either of the magical numbers, 12 or 20. Below is our data analysis.

They're talking about Black women but mention 247 women - that's women of all races shot and killed by police. I can confirm that number. So let's assume they mean women of all races - even though the sentence implies Black women.

```{r}
# Put code to reverse engineer sentence here

#trying with v1 data to look at threat_level and see if we can get at women who were not the culprit

women_fatal_police_shootings_sept_2020 |> group_by(threat_level) |>
  summarise(count = n())
# Display results of code below this codeblock

```

#none of these get close to what we're looking for (20 or 12). Let's try the v2 dataset:

```{r}
women_fatal_police_shootings_2020 |> group_by(threat_type) |>
  summarise(count = n())
```

#Ok not sure what I did differently here but now suddenly accident and flee add to 20. Maybe those are our 20 women? Don't feel super confident about this but not sure where else they would have gotten that number. 

### Sentence 3

* **Sentence text**: At least 89 of the women [fatally shot by police] were at their homes or residences where they sometimes stayed.
* **Analysis summary**: I don't think this was determined through the dataset we have. There's no indicator of this info listed.

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.
* **Analysis summary**: We were able to confirm the findings using the original dataset Joel and Shaun found that reflects what info they had when the article came out. This confirms we should use this dataset going forward.

```{r}
# Put code to reverse engineer sentence here
data_to_sept_2020_shootings |> 
  group_by(gender, signs_of_mental_illness) |>
  summarise(count = n())
# Display results of code below this codeblock
```

This is looking more accurate. 77 out of 247 women and 22% of men had signs of mental illness.
### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-