---
title: "Reverse Engineering Project"
author: "Taylor Joel Shaun & Menna"
date: "7/7/23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been shot by police since 2015](https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
setwd("~/data_journalism/data_journalism_fall_2023/major_assignments/reverse_engineering/Reverse-Engineering-Project/data")

read_csv("fatal-police-shootings-data.csv")
fatal_police_shootings <- read.csv("fatal-police-shootings-data.csv")
women_fatal_police_shootings <- fatal_police_shootings %>%  
  filter(str_detect(gender, "female")) 

count_women_police_shootings <- women_fatal_police_shootings %>% 
  summarize(count = n())

## cleaning the column names so I won't have trouble later on
clean_women_fatal_police_shootings <- women_fatal_police_shootings %>% 
  clean_names()

```

Now, **clean_women_fatal_police_shootings** represents **women_fatal_police_shootings** but cleaned to remove spaces and the like so we can actually work with column names if necessary

Next, we want to clean the "date" column to actually work like a date column with lubridate. To not mess things up with the original data, we're going to create a new variable that represents the above data but only only until 2020, when the Washington Post article was published.

```{r}
clean_2020_women_fatal_police_shootings <- clean_women_fatal_police_shootings

clean_2020_women_fatal_police_shootings
```

Then, we're going to actually transform the column into something useful with lubridate.

```{r}
clean_2020_women_fatal_police_shootings <- clean_2020_women_fatal_police_shootings |> mutate(date=mdy(date))
```

Now, we're finally going to actually transform it into before the publish date.

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "Since The Washington Post began tracking fatal shootings by police in 2015, officers have fatally shot 247 women out of the more than 5,600 people killed overall."
* **Analysis summary**: We were able to get pretty close to the numbers quoted in the article. We came up with 248, one more than the Post included in their article. We aren't sure where the extra person is coming from, but we weren't able to locate them. The data doesn't include timing, so maybe the article came out before one from the same day got logged? That seems like a stretch though.

```{r}
# Put code to reverse engineer sentence here
clean_2020_women_fatal_police_shootings <- clean_2020_women_fatal_police_shootings %>% 
  filter(date<"2020-09-04") %>% 
  summarize(count = n())

clean_2020_women_fatal_police_shootings

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-